---
title: "Influenza A H7N9 in China, 2013"
format: dashboard
theme: morph
---

```{r include=FALSE, echo=FALSE, message=FALSE}
# Load packages 
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, outbreaks, bslib, bsicons, shiny, 
               rgeoboundaries, plotly, htmltools, reactable)

# view data
head(outbreaks::fluH7N9_china_2013)
```

```{r include=FALSE, echo=FALSE, message=FALSE}
data <- outbreaks::fluH7N9_china_2013

n_cases <- data %>% nrow()

n_hospitalization <- data %>% 
  filter(!is.na(date_of_hospitalisation)) %>% 
  nrow()

n_deaths <- data %>% 
  filter(outcome=="Death") %>% 
  nrow()

graph_province <- data %>% 
  group_by(province) %>% 
  count() %>% 
  arrange(-n) %>% 
  mutate(tooltip_label = paste(province, 
                               n, 
                               sep = ": ")) %>% 
  ggplot(mapping = aes(x = n, 
                       y = fct_reorder(province, n),
                       text = tooltip_label)) +
  geom_col(fill = "lightblue") +
  theme_minimal() +
  labs(x = "Number of cases",
       y = "")

graph_province_plotly <- ggplotly(graph_province, tooltip = "text")
```


```{r results='hide', message=FALSE, warning = FALSE}
#create graph
pacman::p_load(outbreaks, rgeoboundaries, plotly)

### MAP OF CASES PER PROVINCE
province_summary <- fluH7N9_china_2013 %>% 
  group_by(province) %>% 
  summarise(total_cases = n()) 

china_map <- rgeoboundaries::gb_adm1("china")

# Check the keys to see if join will be successful
setdiff(province_summary$province, china_map$shapeName)
setdiff(china_map$shapeName, province_summary$province)
# Some work to be done to harmonize these

china_map <- china_map %>%
  # Remove the strings Province, Municipality and Autonomous Region 
  # to allow matching
  mutate(province = str_trim(
    str_remove(shapeName, "Province|Municipality|Autonomous Region|Special Administrative Region")
  ))

# Check if join will work fine
setdiff(province_summary$province, china_map$province)

# Fix Guangdong, which based on a quick Google is actually Guangzhou
china_map <- china_map %>% 
  mutate(province = if_else(province == "Guangzhou", "Guangdong", province)) 
# Check again
setdiff(province_summary$province, china_map$province)

# Now join
china_map_joined <- 
  left_join(china_map, province_summary, by = "province") %>% 
  # Create tooltip label 
  mutate(tooltip_label = paste0(province, ": ", total_cases))

china_map_plot <- ggplot(china_map_joined) +
  geom_sf(aes(fill = total_cases, text = tooltip_label)) + 
  theme_void() +
  labs(fill = "Number of Cases") + 
  theme(legend.position = "none")

china_map_plot_ggplotly <- 
  ggplotly(china_map_plot, tooltip = "text")

```

# HOME 

## {height=20%}

```{r}
value_box(
  title = "Total cases",
  value = n_cases,
  showcase = bsicons::bs_icon("virus"),
  theme = value_box_theme(bg = "#518fd6")
)
```

```{r}
value_box(
  title = "Total hospitalizations",
  value = n_hospitalization,
  showcase = bsicons::bs_icon("hospital"),
  theme = value_box_theme(bg = "#518fd6")
)
```

```{r}
value_box(
  title = "Total deaths",
  value = n_deaths,
  showcase = bsicons::bs_icon("file-medical"),
  theme = value_box_theme(bg = "#518fd6")
)
```

## {height=80%}

### {width=30%}

```{r title = "Number of Cases by Province"}
graph_province_plotly
```

### {width=70%}

```{r title = "Map of Provinces by Number of Cases"}
china_map_plot_ggplotly
```

# DOWNLOAD DATA

```{r}
htmltools::browsable(
  tagList(
    reactable(data, 
              elementId = "data-table", 
              searchable = T, 
              filterable = T), 
    
tags$button("Download as CSV", 
            onclick = "Reactable.downloadDataCSV('data-table')")
  )
)
```

# ABOUT

Avian influenza A(H7N9) is a subtype of influenza viruses that have been detected in birds in the past. This particular A(H7N9) virus had not previously been seen in either animals or people until it was found in March 2013 in China.

However, since then, infections in both humans and birds have been observed. The disease is of concern because most patients have become severely ill. Most of the cases of human infection with this avian H7N9 virus have reported recent exposure to live poultry or potentially contaminated environments, especially markets where live birds have been sold. This virus does not appear to transmit easily from person to person, and sustained human-to-human transmission has not been reported.

For more information you can refer to <https://www.who.int/emergencies/situations/avian-influenza-a-(h7n9)-virus-outbreak>